Locally-Oriented Programming:\\ A Simple Programming Model for Stencil-Based Computations on Multi-Level Distributed Memory Architectures

### 2015-02-07 papers/LOPe/transformations.tex
Paper lenght shortening begun.

\subsection{Halo Transfer Performance}

A thorough examination of potential performance gains using LOPe for parallelization of code is beyond
the scope of this paper.  The emphasis of this work is to explain the syntax and execution symantics
of a LOPe application.  However, preliminary work using the StrategoXT rewrite system as described indicates
that similar performance can be expected to that of previous work\cite{foropencl} (see also\cite{StencilCompilers}).

The current implementation does \emph{not} take advantage of optimization strategies such as
prefetching of array tiles (including halos) into OpenCL local memory or of loop unrolling on the
OpenCL device.  Neither does it take advantage of the potential of LOPe \texttt{CONCURRENT} function
fusion along with execution and synchronization strategies to overlap communication with
computation.

Since many scientific codes are dominated by memory performance, including and especially stencil algorithms
as they typically only involve a computation on a small locally central array element and a small overlapping
halo region.  Stencil operations frequently do not contain enough floating point operations per memory load
to allow for floating point performation to operate at peak (though this is entirely application and domain
specific).  Thus we illustrate the \emph{potential} for performance by noting the latency and throughput
performance of an attached GPU inconjuction with MPI distributed memory performance associated with halo
transfer in Table 1.

The results in Table 1 indicate that the primary bottleneck in using accelerators attached to the
TODO bus using OpenCL is likely to be the latency in transferring memory to and from the device for
distributed memory clusters of only a few nodes exchanging halo data using MPI.

>>>>

%%\subsection{Halo Transfer Performance}

%%A thorough examination of potential performance gains using LOPe for parallelization of code is beyond the scope of this paper.  The emphasis of this work is to explain the syntax and execution symantics of a LOPe application.  However, preliminary work using the StrategoXT rewrite system as described indicates that similar performance can be expected to that of previous work\cite{foropencl} (see also\cite{StencilCompilers}).

%%The current implementation does \emph{not} take advantage of optimization strategies such as prefetching of array tiles (including halos) into OpenCL local memory or of loop unrolling on the OpenCL device.  Neither does it take advantage of the potential of LOPe \texttt{CONCURRENT} function fusion along with execution and synchronization strategies to overlap communication with computation.

%%Since many scientific codes are dominated by memory performance, including and especially stencil algorithms as they typically only involve a computation on a small locally central array element and a small overlapping halo region.  Stencil operations frequently do not contain enough floating point operations per memory load to allow for floating point performation to operate at peak (though this is entirely application and domain specific).  Thus we illustrate the \emph{potential} for performance by noting the latency and throughput performance of an attached GPU inconjuction with MPI distributed memory performance associated with halo transfer in Table 1.

%%The results in Table 1 indicate that the primary bottleneck in using accelerators attached to the TODO bus using OpenCL is likely to be the latency in transferring memory to and from the device for distributed memory clusters of only a few nodes exchanging halo data using MPI.
